### 2015-11-19 data_cleaning.tex
added files

\documentclass{article}
\usepackage{color}
\usepackage{listings}	
\usepackage{enumerate}
\usepackage{layouts}
\usepackage{times}
\usepackage[numbers]{natbib}
\usepackage{notoccite}
\usepackage{url}
\usepackage{xspace}
\usepackage{tikz}

%\author{Authors}

\title{On tabular constraint learning}
\begin{document}
\maketitle
\vspace{-15pt}
\section{Headers}
Can be in columns or in row, see examples. Needs to be inferred using type information.
\vspace{-10pt}
\section{General Algorithm: Infer-Learn-Suggest -- ILS }
\paragraph{Infer} Datatypes, headers, i.e. variable names, and table layouts, borders of tables, the number of tables etc
\vspace{-10pt}
\paragraph{Learn} On each level, i.e., table-wise or column-row-wise learn applicable constraints
\vspace{-10pt}
\paragraph{Suggest} If there is a row or a column with missing data (not-completed cells), then suggest values satisfying the constraints
\vspace{-10pt}
\section*{Types of Constraints}
Most of the constraints I have seen are numeric (finance, accounting) and occasionally strings, also emails and dates are present but their quantity is negligible.

Types of constraints exact or noisy(way harder)
\begin{itemize}
  \item Ascending/Descending (all types)
  \item Ordered by (all types)
  \item All different (all types, similar to the key detection)
  \item Arithmetics (only for number type + possible ranges over two, N, all columns-rows or their consecutive subset)
  \item Concatenation (strings + any other type)
  \item Common substring? 
  \item Column row-inclusion, sort of FD, all values of a row or a column are present in the other row or a column
  \item Aggregates? Counting 
  \item Key -- foreign or primary
\end{itemize}
\end{document}